Application for Rethinking Foundations of Physics 2017

### 2017-01-16 Rethinking Foundations of Physics 2017/application.tex
application.tex Updated...

It is unlikely that there will be a simple a resolution to this
awkward situation. It is more likely that the resolution will emerge
from deep and careful analyses of the foundations of each field. As
the world-renowned logician Girard states, the foundations of physics
and the foundations of logic and computation appear strongly
correlated:
\begin{quote}
  In other terms, what is so good in logic that quantum physics should
  obey? Can't we imagine that our conceptions about logic are wrong,
  so wrong that they are unable to cope with the quantum miracle?
  [\ldots] Instead of teaching logic to nature, it is more reasonable
  to learn from her. Instead of interpreting quantum into logic, we
  shall interpret logic into quantum~\cite{MSC:1444836}.
\end{quote}

@@

In the kick-off talk, we will begin by reviewing existing work that
recasts classical probability spaces in a resource-aware setting and
move to aim at recasting quantum probability and quantum measurement.
In particular, we develop a measurement framework based on quantum
interval-valued probability measures~(IVPM). Surprisingly, we found
a quantum IVPM which can not be induced by a state, while Shapley
proved a classical convex IVPM can always be induced by a classical
``state''~\cite{Shapley1971,Grabisch2016}, and Gleason proved
a infinitely-specified quantum probabilities measure can always be
induced by a quantum state~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.
Nevertheless, a quantum IVPM could still correspond to many possible
states if it is broken into pieces. Our examples also suggest that
a quantum IVPM could be more and more likely induced by a state while
the intervals become sharper, i.e., the measurement resources increase.

@@

\item For any event $E$, $\bar{\mu}\left(\Omega\backslash E\right)=\left[1,1\right]-\bar{\mu}\left(E\right)$ 
\item For a collection $\left\{ E_{i}\right\} _{i=1}^{M}$ of pairwise disjoint
events, we have $\bar{\mu}\left(\bigcup_{i=1}^{M}E_{i}\right)\subseteq\sum_{i=1}^{M}\bar{\mu}\left(E_{i}\right)$. 

@@

\begin{example}[Two-coin experiment with interval probability]\label{ex3}

@@

\begin{example}[Three-dimensional quantum four-interval-valued probability measure]\label{ex:three-dimensional-4-value}
Given a three dimensional Hilbert space with an orthonormal basis
$\left\{ \ket{0},\ket{1},\ket{2}\right\} $. Let \emph{ }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
$\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, and $\ket{\ms}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})$.
The definition of $\bar{\mu}$ below refers to Fig.~\ref{fig:three-dimensional-4-value}
which plots the 1-dimensional projectors: 
\begin{enumerate}
\item Let: 
\begin{eqnarray*}
\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps}) & = & \imposs\\
\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps}) & = & \necess
\end{eqnarray*}
where $\ket{0}$ and $\ket{\ps}$ are plotted as the red and green
dotted vectors, respectively. 
\item The red and green circles are are the states orthogonal to $\ket{0}$
and $\ket{\ps}$, respectively, and can be parametrized as $\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
and $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\ms}+\cos\theta\ket{2}$,
where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. The dotted
half of those states need special treatment, i.e., whenever $0\le\theta<\frac{\pi}{2}$
and $0\le\gamma<\pi$, we define 
\begin{eqnarray*}
\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\proj{\ps_{\theta,\gamma}^{\perp}}\right) & = & \likely\\
\bar{\mu}\left(\mathbb{1}-\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\mathbb{1}-\proj{\ps_{\theta,\gamma}^{\perp}}\right) & = & \unlikely
\end{eqnarray*}
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\likely$. 
\end{enumerate}
It is straightforward but tedious to check that $\bar{\mu}$ is a
quantum interval-valued probability measure. Although we have not
verified that $\bar{\mu}$ is convex, the following argument establishes
that is has an empty core. Assume there is a real-valued probability
measure satisfying $\mu_{\rho}^{B}(P)\in\bar{\mu}(P)$ for all $P\in\events$.
Because $\mu_{\rho}^{B}(\proj{0})\in\bar{\mu}(\proj{0})=\imposs$
and $\mu_{\rho}^{B}(\proj{\ps})\in\bar{\mu}(\proj{\ps})=\imposs$,
we must have $\mu_{\rho}^{B}(\proj{0})=\mu_{\rho}^{B}(\proj{\ps})=0$
so that $\mu_{\rho}^{B}=\mu_{\ket{2}}^{B}$. However, 
\[
\mu_{\ket{2}}^{B}(\proj{2})=1\notin\unlikely=\bar{\mu}(\proj{2})\textrm{ .}
\]
This measure however is ``better'' than the previous one in the
sense that it might only be induced by $\ket{2}$ or the density matrix
$\frac{\mathbb{1}}{3}$, but not by $\ket{1}$. \end{example}

@@

Although a family of more and more precise quantum IVPMs might be
induced from a real quantum state, given a particular quantum IVPM,
i.e., using a particular measurement apparatus, our investigation
leaves an open question of whether behind it there exists a ``real''
state. Despite the possibility to use quantum IVPMs as the most primitive
definition of states, we might focus on the meaning of the probability.
Since a quantum IVPM cannot be defined without consider a particular
measurement apparatus, and the information extracted from this apparatus,
the interval-valued probability itself could be considered subjective,
and could be interpreted in the Bayesian sense as ``lack of knowledge''.
This, and similar arguments elegantly expressed by recent work on
Quantum Bayesianism~\cite{Fuchs2010,VonBaeyer2016,Fuchs2012}, which
suggests that the quantum state is more like an interactive system
in computer science parlance. In other words, the state is only determined
by its interactions with the environment or observers. A rough but
useful analogy in the context of computer science would be a web server:
the state of such a system cannot be said to exist independently of
the interactions with clients; each interaction with a client determines
a future state of the system and may affect future interactions. Crucially,
each client may have a \emph{different} view of the web server that
is consistent with their own history of interactions. This computational
perspective agrees with the philosophical position of QBism~\cite{VonBaeyer2016}
that asserts that the quantum state is subjective: each observer has
a different view of the quantum system that is consistent with their
previous observations and that allows that observer, independently
of other observers, to assign beliefs (i.e., probabilities) to possible
future interactions.

@@

I would like to apply for funding, but I haven’t been to Europe before…
So I don't know the specific amount I need…

>>>>

It is unlikely that there will be a simple resolution to this
awkward situation. It is more likely that the resolution will emerge
from deep and careful analyses of the foundations of each field. As
the world-renowned logician Girard states, the foundations of physics
and the foundations of logic and computation appear strongly
correlated:
\begin{quote}
  In other terms, what is so good in logic that quantum physics should
  obey? Can't we imagine that our conceptions about logic are wrong,
  so wrong that they are unable to cope with the quantum miracle?
  [\ldots] Instead of teaching logic to nature, it is more reasonable
  to learn from her. Instead of interpreting quantum into logic, we
  shall interpret logic into quantum~\cite{MSC:1444836}.
\end{quote}

@@

In the kick-off talk, we will begin by reviewing existing work that
recasts classical probability spaces in a resource-aware setting and
move to aim at recasting quantum probability and quantum measurement.
In particular, we develop a measurement framework based on quantum
interval-valued probability measures. Surprisingly, we found a quantum
IVPM which can not be induced by a state, while Shapley proved a classical
convex IVPM can always be induced by a classical ``state''~\cite{Shapley1971,Grabisch2016},
and Gleason proved a infinitely-specified quantum probability measure
can always be induced by a quantum state~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.
Nevertheless, a quantum IVPM could still correspond to many possible
states if it is broken into pieces. Our examples also suggest that
a quantum IVPM could be more and more likely induced by one state
while the intervals become sharper, i.e., the measurement resources
increase.

@@

\item For any event $E$, $\bar{\mu}\left(\Omega\backslash E\right)=\left[1,1\right]-\bar{\mu}\left(E\right)$.
\item For a collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise disjoint
events, we have $\bar{\mu}\left(\bigcup_{i=1}^{N}E_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(E_{i}\right)$. 

@@

\begin{example}{[}Two-coin experiment with interval probability{]}\label{ex3}

@@

\begin{example}[Three-dimensional quantum four-interval-valued probability measure]\label{ex:three-dimensional-4-value}
Given a three dimensional Hilbert space with an orthonormal basis
$\left\{ \ket{0},\ket{1},\ket{2}\right\} $. Let \emph{ }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
$\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, and $\ket{\ms}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})$.
The definition of $\bar{\mu}$ below refers to Fig.~\ref{fig:three-dimensional-4-value}
which plots the 1-dimensional projectors: 
\begin{enumerate}
\item Let: 
\begin{eqnarray*}
\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps}) & = & \imposs\\
\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps}) & = & \necess
\end{eqnarray*}
where $\ket{0}$ and $\ket{\ps}$ are plotted as the red and green
dotted vectors, respectively. 
\item The red and green circles are the states orthogonal to $\ket{0}$
and $\ket{\ps}$, respectively, and can be parametrized as $\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
and $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\ms}+\cos\theta\ket{2}$,
where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. The dotted
half of those states need special treatment, i.e., whenever $0\le\theta<\frac{\pi}{2}$
and $0\le\gamma<\pi$, we define 
\begin{eqnarray*}
\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\proj{\ps_{\theta,\gamma}^{\perp}}\right) & = & \likely\\
\bar{\mu}\left(\mathbb{1}-\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\mathbb{1}-\proj{\ps_{\theta,\gamma}^{\perp}}\right) & = & \unlikely
\end{eqnarray*}
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\likely$. 
\end{enumerate}
It is straightforward but tedious to check that $\bar{\mu}$ is a
quantum interval-valued probability measure. Although we have not
verified that $\bar{\mu}$ is convex, the following argument establishes
that is has an empty core. Assume there is a real-valued probability
measure satisfying $\mu_{\rho}^{B}(P)\in\bar{\mu}(P)$ for all $P\in\events$.
Because $\mu_{\rho}^{B}(\proj{0})\in\bar{\mu}(\proj{0})=\imposs$
and $\mu_{\rho}^{B}(\proj{\ps})\in\bar{\mu}(\proj{\ps})=\imposs$,
we must have $\mu_{\rho}^{B}(\proj{0})=\mu_{\rho}^{B}(\proj{\ps})=0$
so that $\mu_{\rho}^{B}=\mu_{\ket{2}}^{B}$. However, 
\[
\mu_{\ket{2}}^{B}(\proj{2})=1\notin\unlikely=\bar{\mu}(\proj{2})\textrm{ .}
\]
This measure however is ``better'' than the previous one in the
sense that it might only be induced by $\ket{2}$ or the density matrix
$\frac{\mathbb{1}}{3}$, but not by $\ket{1}$. \end{example}

@@

Although a family of more and more precise quantum IVPMs might be
induced from a real quantum state, given a particular quantum IVPM,
i.e., using a particular measurement apparatus, our investigation
leaves an open question of whether behind it there exists a ``real''
state. Despite the possibility of using quantum IVPMs as the most
primitive definition of states, another possibility is to focus on
the meaning of the probability.

Since a quantum IVPM cannot be defined without considering a particular
measurement apparatus, the interval-valued probability itself could
be considered subjective, and could be interpreted in the Bayesian
sense as ``lack of knowledge''. This, and similar arguments elegantly
expressed by recent work on Quantum Bayesianism~\cite{Fuchs2010,VonBaeyer2016,Fuchs2012},
which suggests that the quantum state is more like an interactive
system in computer science parlance. In other words, the state is
only determined by its interactions with the environment or observers.
A rough but useful analogy in the context of computer science would
be a web server: the state of such a system cannot be said to exist
independently of the interactions with clients; each interaction with
a client determines a future state of the system and may affect future
interactions. Crucially, each client may have a \emph{different} view
of the web server that is consistent with their own history of interactions.
This computational perspective agrees with the philosophical position
of QBism~\cite{VonBaeyer2016} that asserts that the quantum state
is subjective: each observer has a different view of the quantum system
that is consistent with their previous observations and that allows
that observer, independently of other observers, to assign beliefs
(i.e., probabilities) to possible future interactions.

@@

Yes.
